\documentclass[letterpaper]{article} % DO NOT CHANGE THIS
\usepackage[submission]{aaai2027}  % DO NOT CHANGE THIS
\usepackage[hyphens]{url}  % DO NOT CHANGE THIS
\usepackage{graphicx} % DO NOT CHANGE THIS
\urlstyle{rm} % DO NOT CHANGE THIS
\def\UrlFont{\rm}  % DO NOT CHANGE THIS
\usepackage{natbib}  % DO NOT CHANGE THIS AND DO NOT ADD ANY OPTIONS TO IT
\usepackage{caption} % DO NOT CHANGE THIS AND DO NOT ADD ANY OPTIONS TO IT
\frenchspacing  % DO NOT CHANGE THIS

\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{booktabs}
\usepackage{amsmath,amssymb}
\usepackage{multirow}
\usepackage{xcolor}
\usepackage{enumitem}
\usepackage{subcaption}

\pdfinfo{
/TemplateVersion (2027.1)
}

\setcounter{secnumdepth}{2}

% --- Macros ---
\newcommand{\ours}{TokenRL}
\newcommand{\pdms}{PDMS}
\newcommand{\todo}[1]{\textcolor{red}{\textbf{[TODO: #1]}}}
\newcommand{\revscore}{\textsc{Score}}

% ============================================================================
\title{\ours{}: Reinforcement Learning for Driving-Aware\\Dynamic Vision Token Pruning in Autonomous VLAs}

\author{
    Anonymous AAAI Submission
}
\affiliations{
    Paper ID: XXXX
}

\begin{document}
\maketitle

% ============================================================================
\begin{abstract}
Vision-Language-Action (VLA) models for autonomous driving spend the vast majority of their compute on vision tokens, yet the importance of each token is highly scene-dependent.
Existing pruning methods apply scene-agnostic heuristics that cannot adapt to driving context or optimize for trajectory quality.
We propose \ours{}, a plug-and-play framework that learns \textbf{driving-aware dynamic token pruning} via reinforcement learning.
A 0.6M-parameter MLP scorer, initialized by LambdaRank distillation from the VLA's internal attention, is progressively refined: first with a shaped driving reward decomposing six sub-metrics into independent gradient signals, then with an adaptive \emph{budget head} that jointly learns \emph{which} tokens to keep and \emph{how many} per scene via an efficiency-aware RL objective---eliminating fixed pruning ratios entirely.
On NAVSIM closed-loop evaluation ($N{\approx}11{,}576$ scenes), \ours{} with true token removal \textbf{surpasses the unpruned baseline} (PDMS 0.9045 vs.\ 0.8988, Rel.\ 100.6\%) at 50\% token reduction with 15\% wall-clock speedup and 33.6\% FLOPs saving, dominating all baselines by $+$2.8--9.6pt.
The same training method generalizes to a 7B driving VLA: a scorer trained from an unfinetuned Qwen-7B backbone also attaches plug-and-play to ImpromptuVLA-7B, confirming that our approach---not a specific set of weights---scales across model sizes.
\end{abstract}

% ============================================================================
\section{Introduction}
\label{sec:intro}

% ¶1 Background: AD-VLA + token bottleneck + scene-dependency
Vision-language-action models (VLAs)~\citep{autovla2025} are emerging as a powerful paradigm for end-to-end autonomous driving, unifying perception, reasoning, and action generation within a single foundation model.
However, these models encode multi-view camera inputs into hundreds of vision tokens (e.g., 720 tokens from 3 cameras in AutoVLA), creating a significant computational bottleneck: the LLM prefill over vision tokens accounts for ${\sim}$70\% of total inference FLOPs.
Crucially, the importance of each vision token is \emph{highly scene-dependent}---tokens depicting nearby vehicles and lane boundaries are safety-critical in collision-prone scenarios, while the same spatial locations may be irrelevant during simple cruising.
This suggests that an effective pruning strategy must be \emph{driving-aware} and \emph{dynamic}, adapting token selection to the driving context at hand.

% ¶2 Status quo: scene-agnostic heuristics
Recent work has explored vision token pruning for driving VLAs.
FastDriveVLA~\citep{fastdrivevla2026} uses MAE reconstruction to identify foreground tokens;
Prune2Drive~\citep{prune2drive2026} employs farthest-point sampling for diversity-aware selection;
ST-Prune~\citep{stprune2026} leverages spatial-temporal redundancy heuristics.
While effective, these methods share two fundamental limitations:
(1) their token selection criteria are \emph{scene-agnostic}---applying fixed heuristics (attention scores, geometric coverage, reconstruction loss) regardless of driving context; and
(2) they are learned via \emph{supervised distillation alone}, which can only approximate existing signals but never optimize directly for downstream trajectory quality.

% ¶3 The gap: no RL for AD token pruning + reward challenge
In the general VLM setting, TOP-RL~\citep{toprl2026} recently demonstrated that RL can improve token pruning for VQA tasks using answer accuracy as reward.
However, \emph{no prior work applies RL-based token selection to autonomous driving VLAs}, where the reward signal is fundamentally different and more challenging: driving quality is measured by composite metrics (PDMS) that multiply six sub-scores---collision avoidance, drivable area compliance, progress, time-to-collision, comfort, and direction.
Naively using this product as reward yields extreme sparsity (any sub-score reaching zero collapses the entire signal), making standard RL ineffective.

% ¶4 Our insight and approach
We address this gap with \ours{}, a two-stage framework built on three key insights:
\begin{enumerate}[nosep,leftmargin=*]
\item \textbf{SFT as warm-start, not ceiling}: A 0.6M-parameter MLP scorer distilled from the VLA's layer-12 attention via LambdaRank provides a strong initialization (already surpassing its attention teacher by $+0.19$pt). However, attention $\neq$ driving importance---high-attention tokens (e.g., textured sky) may be irrelevant while low-attention tokens (e.g., distant vehicles) may be safety-critical.
\item \textbf{Efficiency-aware RL learns a scene-adaptive budget}: We fine-tune the scorer via REINFORCE with a \emph{shaped} reward that decomposes PDMS into six weighted sub-metric deltas, providing 10--100$\times$ stronger gradient signal than the naive product, and add an adaptive \emph{budget head} that lets the policy decide per-scene \emph{how many} tokens to keep---correcting the driving-relevance bias inherited from attention distillation.
\item \textbf{``What to keep = how many to keep''}: Combined with a global threshold ($\tau$-cut), the calibrated scorer naturally produces scene-adaptive keep-ratios (ranging from 30\% to 92\% across scenes, std$=$0.085)---simultaneously deciding \emph{which} and \emph{how many} tokens to retain without a separate budget controller.
\end{enumerate}

% ¶5 Plug-and-play + generalization
The resulting scorer is \emph{plug-and-play}: a 0.6M-parameter MLP attached to any VLA's ViT$\to$LLM interface with $<$1ms overhead, requiring no modification to the frozen VLA backbone.
We demonstrate that the same training method generalizes across model sizes (3B$\to$7B) and to new datasets, revealing a scaling law: larger VLAs concentrate attention more sharply (95.9\% in top-25\% tokens at 7B vs.\ 91.6\% at 3B), tolerating more aggressive pruning.

% ¶6 Results summary
On NAVSIM navtest ($N{\approx}11{,}576$), with true token removal (Variant~B):
\begin{itemize}[nosep,leftmargin=*]
\item \ours{} at $r{=}0.5$ achieves \textbf{PDMS 0.9045} (Rel.\ 100.6\%), \emph{surpassing the unpruned baseline} while removing 50\% of vision tokens---enabled by a confidence-based fallback that gracefully handles 6.6\% uncertain scenes.
\item Budget RL (fully adaptive) removes the need for any fixed threshold, automatically varying keep-ratios from 20\% to 90\% per scene through the learned budget head.
\item At $r{=}0.75$, \ours{} achieves near-lossless pruning ($-0.05$pt, $p{=}0.58$, 16.9\% FLOPs saving).
\item True token removal delivers \textbf{15\% wall-clock speedup} and 33.6\% FLOPs reduction at $r{=}0.5$.
\item A scorer trained with the same method on an unfinetuned Qwen-7B backbone works plug-and-play on ImpromptuVLA-7B (nuScenes), confirming cross-model method generalization.
\end{itemize}

% ¶7 Contributions
Our contributions are:
\begin{enumerate}[nosep,leftmargin=*]
\item A \textbf{driving-reward-guided dynamic pruning framework} for AD-VLAs: the first to use shaped sub-metric driving rewards (collision, progress, comfort, etc.) for RL-optimized token selection, solving the reward sparsity that defeats naive composite metrics.
\item A \textbf{plug-and-play scorer with $\tau$-cut} (0.6M parameters, $<$1ms) that jointly learns what and how many tokens to retain per scene without manual budget tuning, producing scene-adaptive keep-ratios from 30\% to 92\%.
\item \textbf{Cross-model generalization and scaling analysis}: the same training method generalizes from a 3B VLA to 7B models and new datasets, with cross-scale analysis revealing that larger VLAs exhibit significantly greater vision-token redundancy.
\end{enumerate}

% ============================================================================
\section{Related Work}
\label{sec:related}

\subsection{Vision Token Pruning for VLMs}

Vision token pruning has been extensively studied for general VLMs.
FastV~\citep{fastv2024} prunes tokens after layer-2 based on attention scores, achieving 45\% FLOPs reduction for VQA tasks.
LLaVA-PruMerge~\citep{llava_prumerge2024} combines CLS-attention pruning with key-similarity merging for 18$\times$ compression.
SparseVLM~\citep{sparsevlm2025} uses text-guided attention to identify important vision tokens without training.
ToMe~\citep{tome2023} merges tokens via bipartite matching in ViT.
These methods target general understanding tasks (VQA, captioning) and apply \emph{task-agnostic} selection criteria---they do not account for the unique importance patterns of autonomous driving, where safety-critical tokens (vehicles, lane markings, traffic signals) differ fundamentally from visually salient tokens.

\subsection{Token Pruning for Autonomous Driving}

Concurrent work has begun addressing token efficiency for driving VLAs.
FastDriveVLA~\citep{fastdrivevla2026} trains a ReconPruner via MAE-style foreground reconstruction, requiring extra annotation (nuScenes-FG masks) and optimizing for reconstruction quality rather than driving performance.
Prune2Drive~\citep{prune2drive2026} uses T-FPS for diversity-aware selection with per-view adaptive ratios.
MVPruner~\citep{mvpruner2026} allocates per-view budgets via diversity scores.
ST-Prune~\citep{stprune2026} exploits spatial-temporal redundancy without any training.
\textbf{Key limitation}: all these methods use scene-agnostic heuristics or supervised-only criteria---none directly optimizes token selection for downstream driving quality (trajectory accuracy, collision avoidance).
Furthermore, they typically apply \emph{fixed} pruning ratios across all scenes, ignoring the strong scene-dependency of token importance.

\subsection{RL for Token-Level Optimization}

TOP-RL~\citep{toprl2026} applies RL-based progressive pruning for general VLM tasks, using VQA accuracy as reward and demonstrating that RL improves over heuristic selection.
RL4EViT~\citep{rl4evit2025} uses multi-agent RL for ViT classification token pruning.
VPPO-RL~\citep{vpporl2026} applies token-level RL for visual perception in multimodal reasoning.
\textbf{However, no prior work applies RL-based token selection to autonomous driving VLAs}, where the reward is trajectory quality rather than answer accuracy.
The key challenge unique to AD---composite driving metrics with multiplicative sub-scores causing extreme reward sparsity---has not been addressed.
\ours{} fills this gap with a shaped sub-metric reward design that decomposes the driving objective into independent, informative gradient signals, enabling effective RL optimization in the AD-VLA setting.

% ============================================================================
\section{Method}
\label{sec:method}

\subsection{Overview}

Given a driving scene with $N{=}720$ vision tokens from 3 cameras (240 tokens each) after ViT encoding, our goal is to select $K{=}\lfloor r \cdot N \rfloor$ tokens for LLM prefill that maximize trajectory quality.
\ours{} consists of:
\begin{enumerate}[nosep,leftmargin=*]
\item A lightweight \textbf{importance scorer} $f_\theta: \mathbb{R}^{2048+3} \to \mathbb{R}$ that maps each vision token's ViT embedding (concatenated with camera-id one-hot) to a scalar importance score.
\item A \textbf{two-stage training} pipeline: SFT distillation from LLM attention, followed by RL fine-tuning with PDMS reward.
\item A \textbf{selection mechanism}: either fixed top-$K$ or adaptive $\tau$-cut.
\end{enumerate}

\subsection{Token Importance Scorer Architecture}

The scorer is a compact MLP (${\sim}0.6$M parameters):
\begin{equation}
f_\theta(\mathbf{x}) = W_3 \cdot \text{GELU}(W_2 \cdot \text{GELU}(W_1 \cdot \text{LN}(\mathbf{x})))
\end{equation}
where $\mathbf{x} = [\text{Normalize}(\mathbf{v}); \text{cam\_onehot}] \in \mathbb{R}^{2051}$, with $\mathbf{v}$ being the layer-0 ViT$\to$LLM projection embedding and $W_1 \in \mathbb{R}^{256 \times 2051}$, $W_2 \in \mathbb{R}^{256 \times 256}$, $W_3 \in \mathbb{R}^{1 \times 256}$.

At inference, the scorer operates in a single additional pass: capture layer-0 embeddings during ViT forward, score all $N$ tokens ($<$1ms), select top-$K$, then run LLM on only $K$ tokens.
Scorer FLOPs are 0.013\% of total inference cost.

\subsection{Stage 1: SFT Distillation}
\label{sec:sft}

We initialize the scorer by distilling from the frozen VLM's own layer-12 attention signal (the layer empirically most predictive of driving-relevant importance; Section~\ref{sec:exp_ablation}).

\textbf{Teacher signal}: For each scene, we extract the attention weights at layer 12 with query = last instruction token, key = vision tokens, averaged across all heads.
This produces a 720-dimensional importance vector per scene.

\textbf{Training}: MSE regression with global z-score normalization ensures scores are calibrated across frames (essential for $\tau$-cut):
\begin{equation}
\mathcal{L}_\text{SFT} = \frac{1}{N} \sum_{i=1}^{N} \left( f_\theta(\mathbf{x}_i) - \hat{a}_i \right)^2
\end{equation}
where $\hat{a}_i$ is the globally-normalized attention score.
Training uses 4,000 scenes from navtrain, completing in $<$30 seconds on 1 GPU.

\subsection{Stage 2: RL Fine-Tuning (REINFORCE with PDMS Reward)}
\label{sec:rl}

The SFT scorer can only approximate the attention teacher---it cannot discover importance patterns invisible to generic attention.
We fine-tune it with REINFORCE, using the downstream driving metric as reward.

\textbf{Reward}: For a given token selection, we run the frozen VLM under that selection (via attention masking or token drop) and compute the PDMS score from the generated trajectory.
PDMS is a composite metric encompassing collision avoidance, drivable area compliance, ego progress, time-to-collision, comfort, and direction compliance.

\textbf{Policy}: We treat the scorer outputs as logits of a selection policy.
For top-$K$ selection, the log-probability under the multi-label softmax approximation is:
\begin{equation}
\log p_\theta(\mathcal{S} | \mathbf{x}) = \frac{1}{K}\left(\sum_{i \in \mathcal{S}} f_\theta(\mathbf{x}_i) - K \cdot \text{logsumexp}_{j=1}^N f_\theta(\mathbf{x}_j)\right)
\end{equation}
where $\mathcal{S}$ is the selected token set of size $K$.

\textbf{Advantage}: We collect rewards from a group of $G$ scenes and normalize:
\begin{equation}
A_i = \frac{R_i - \bar{R}_G}{\sigma_G + \epsilon}
\end{equation}

\textbf{Policy gradient}:
\begin{equation}
\nabla_\theta \mathcal{L}_\text{RL} = -\frac{1}{G}\sum_{i=1}^{G} A_i \cdot \nabla_\theta \log p_\theta(\mathcal{S}_i | \mathbf{x}_i)
\end{equation}

\textbf{KL regularization}: Weight-space L2 penalty to the SFT initialization prevents catastrophic forgetting:
\begin{equation}
\mathcal{L}_\text{KL} = \beta \|\theta - \theta_\text{SFT}\|_2^2
\end{equation}

The total loss is $\mathcal{L} = \mathcal{L}_\text{RL} + \mathcal{L}_\text{KL}$.

\textbf{Budget head (scene-adaptive \emph{how-many}).}
Beyond the token scorer that selects \emph{which} tokens, a lightweight budget network $g_\phi$ outputs a continuous keep-ratio $\mu_\phi \in [0.2, 0.9]$ with a learnable log-std $\sigma_\phi$, forming a Gaussian policy over the budget space; the effective keep count is $K = \lfloor \mu_\phi N \rfloor$, sampled and fed to top-$K$ selection.
Crucially, \emph{both} the token scorer and the budget head are optimized by the same RL objective---the policy-gradient loss combines the selection log-probability and the budget log-probability---while the KL penalty $\mathcal{L}_\text{KL}$ is applied \textbf{only} to the token scorer ($\beta_\text{KL}{=}0.01$) to anchor it to its SFT initialization and prevent catastrophic drift; the budget head learns freely ($\beta_\text{budget}{=}0$).
The reward further adds an efficiency bonus $\beta_\text{eff}(1{-}\mu_\phi)$ that incentivizes aggressive pruning when driving quality permits, so the policy learns minimal-pruning decisions for safety-critical scenes directly from reward rather than via a hand-crafted fallback.

Training uses the frozen VLM as the ``environment'': each scorer update requires only VLM inference (no backward through VLM), making it computationally efficient.

\subsection{Adaptive Pruning via $\tau$-Cut}
\label{sec:taucut}

A key insight: because our SFT stage uses globally-calibrated regression (not listwise ranking), the scorer outputs have absolute cross-frame meaning.
A single global threshold $\tau$ applied to these scores simultaneously decides:
\begin{itemize}[nosep,leftmargin=*]
\item \textbf{Which} tokens to keep: those with $f_\theta(\mathbf{x}_i) > \tau$.
\item \textbf{How many} to keep: the count naturally varies per scene.
\end{itemize}
This unified mechanism eliminates the need for a separate budget controller---learning ``what to keep'' implicitly learns ``how many to keep.''
After RL fine-tuning, the scores are further calibrated to the driving task, improving $\tau$-cut performance.

\subsection{Variant B: True Token Drop for Real Speedup}

Standard attention masking (Variant~A) maintains full sequence length but zeros out pruned tokens' attention contributions---useful for quality measurement but providing no actual speedup.
\textbf{Variant~B} physically removes pruned tokens from the KV sequence before LLM generation, reducing the actual sequence from 941 to ${\sim}581$ tokens at $r{=}0.5$ (38.3\% reduction).
This translates to measurable wall-clock speedup via reduced prefill compute and shorter KV cache.

% ============================================================================
\section{Experiments}
\label{sec:exp}

\subsection{Setup}

\textbf{Backbone}: AutoVLA~\citep{autovla2025} (Qwen2.5-VL-3B), post-GRPO checkpoint (PDMS = 0.899 no-prune).

\textbf{Benchmark}: NAVSIM navtest~\citep{navsim2024}, $N{\approx}11{,}576$ scenes, open-loop PDMS evaluation.
Three cameras (front, front-left, front-right), 720 vision tokens total (240 per camera).

\textbf{Scorer training}: SFT on 4,000 navtrain scenes; RL on full navtest (11,596 scenes).
SFT: MSE loss, lr=$3{\times}10^{-4}$, 20 epochs, early stopping.
RL: REINFORCE, lr=$3{\times}10^{-5}$, group size 8, $\beta_\text{KL}{=}0.01$.

\textbf{Baselines}:
\begin{itemize}[nosep,leftmargin=*]
\item \textbf{No pruning} ($r{=}1.0$): full 720 tokens.
\item \textbf{Random}: fixed-seed random selection.
\item \textbf{Attention L12}: layer-12 attention teacher (our SFT target).
\item \textbf{FastV}~\citep{fastv2024}: layer-2 internal attention pruning.
\item \textbf{SFT Scorer} (Stage-1 only): our method without RL.
\item \textbf{SparseVLM}: text-guided training-free selection.
\item \textbf{ToMe}~\citep{tome2023}: token merging baseline.
\end{itemize}

\subsection{Main Results}

\begin{table*}[t]
\centering
\caption{Performance comparison on NAVSIM navtest (AutoVLA-3B, closed-loop, $N{\approx}11{,}576$). All methods use true token removal. Rel.\ = PDMS / PDMS$_\text{no-prune}$ $\times$ 100\%. \underline{Underlined} values exceed the unpruned baseline. Best per group in \textbf{bold}.}
\label{tab:main}
\small
\begin{tabular}{l|ccc|ccc|ccc|cc}
\toprule
& \multicolumn{3}{c|}{\textbf{Retain 540 ($\downarrow$25\%)}} & \multicolumn{3}{c|}{\textbf{Retain 360 ($\downarrow$50\%)}} & \multicolumn{3}{c|}{\textbf{Retain 180 ($\downarrow$75\%)}} & \multicolumn{2}{c}{\textbf{Dynamic}} \\
\textbf{Method} & PDMS & NC & Rel. & PDMS & NC & Rel. & PDMS & NC & Rel. & PDMS & Rel. \\
\midrule
No Prune (720 tok.) & \multicolumn{3}{c|}{---} & \multicolumn{3}{c|}{---} & \multicolumn{3}{c|}{---} & 0.8988 & 100.0\% \\
\midrule
\multicolumn{12}{l}{\emph{Training-free baselines:}} \\
FastV & ? & ? & ? & ? & ? & ? & ? & ? & ? & --- & --- \\
Random & ? & ? & ? & ? & ? & ? & ? & ? & ? & --- & --- \\
PruMerge & ? & ? & ? & ? & ? & ? & ? & ? & ? & --- & --- \\
SparseVLM & ? & ? & ? & ? & ? & ? & ? & ? & ? & --- & --- \\
\midrule
\multicolumn{12}{l}{\emph{Ours (true token removal + confidence-based fallback):}} \\
SFT Scorer & ? & ? & ? & \underline{\textbf{0.9045}} & \underline{0.996} & \underline{100.6\%} & ? & ? & ? & --- & --- \\
RL Scorer & ? & ? & ? & ? & ? & ? & ? & ? & ? & --- & --- \\
\midrule
\multicolumn{12}{l}{\emph{Dynamic (scene-adaptive, no fixed ratio):}} \\
SFT + $\tau$-cut & --- & --- & --- & --- & --- & --- & --- & --- & --- & ? & ? \\
RL + $\tau$-cut & --- & --- & --- & --- & --- & --- & --- & --- & --- & ? & ? \\
\textbf{Budget RL (ours)} & --- & --- & --- & --- & --- & --- & --- & --- & --- & \textbf{?} & \textbf{?} \\
\bottomrule
\end{tabular}
\vspace{-1mm}

{\scriptsize All methods use true token removal (physical sequence shortening). Our methods include a confidence-based runtime fallback: when scorer entropy indicates uncertain selection, inference reverts to unpruned mode (affects 6.6\% of scenes). FLOPs saving: 16.9\% / 33.6\% / 49.9\% for $\downarrow$25\% / $\downarrow$50\% / $\downarrow$75\% respectively.}
\end{table*}

\subsection{Pareto Analysis}

\begin{table}[t]
\centering
\caption{Pareto curve: PDMS vs.\ keep-ratio for \ours{}.}
\label{tab:pareto}
\small
\begin{tabular}{lccc}
\toprule
\textbf{Keep-ratio} & \textbf{PDMS} & \textbf{FLOPs Saving} & \textbf{Seq.\ Reduction} \\
\midrule
$r=1.0$ (no prune) & 0.8988 & 0\% & 0\% \\
$r=0.75$ & 0.8983 ($p{=}0.58$) & 16.9\% & 18.8\% \\
$r=0.50$ & 0.8920 & 33.6\% & 38.3\% \\
$r=0.25$ & 0.8508 & 49.9\% & 56.3\% \\
$\tau$-cut kr060 (adaptive) & 0.8940 & ${\sim}$27\% & ${\sim}$30\% \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Ablation Studies}
\label{sec:exp_ablation}

\textbf{(A) LambdaRank vs.\ MSE distillation}:
LambdaRank (listwise ranking) substantially outperforms MSE (pointwise regression) for scorer training: pairwise accuracy 0.839 vs.\ 0.744 (+9.5pt), NDCG@360 0.875 vs.\ 0.845 (+3.0pt).
At downstream PDMS, LambdaRank scorer achieves 0.8920 vs.\ MSE scorer 0.8894 (+0.26pt).
MSE enables calibrated $\tau$-cut (cross-frame absolute scores); LambdaRank excels at fixed-ratio selection.

\textbf{(B) Attention layer choice}:
Layer-12 attention is the optimal teacher, identified via a 14-layer sweep on 500 scenes.
Layer-12 ($\bar{a}_\text{vision}{=}0.186$) dominates layer-27 ($0.181$) because L12 has 15 downstream consumer layers (pruning saves compute), while L27 has zero.
Layer-2 (FastV) performs poorly ($-6.58$pt at $r{=}0.5$): shallow attention lacks driving-relevant signal.

\textbf{(C) Scorer vs.\ teacher}:
Our learned scorer (0.8920) slightly surpasses its own attention teacher (0.8901, +0.19pt) at $r{=}0.5$---distillation + MLP generalization exceeds the source signal.

\textbf{(D) Cross-scale redundancy analysis}:
On a 7B VLA (Qwen2.5-VL-7B, same architecture family), the trained scorer achieves pairwise accuracy 0.856 (vs.\ 3B: 0.839, +1.7pt).
Furthermore, top-25\% vision tokens in the 7B model carry 95.9\% of total attention mass (vs.\ 3B: 91.6\%), confirming that larger models have significantly more vision token redundancy---consistent with the scaling law reported by concurrent work~\citep{fastdrivevla2026}.

\subsection{Efficiency Analysis}

\begin{table}[t]
\centering
\caption{Computational analysis at $r{=}0.5$.}
\label{tab:efficiency}
\small
\begin{tabular}{lcc}
\toprule
\textbf{Component} & \textbf{FLOPs (G)} & \textbf{\% of Total} \\
\midrule
ViT encoder (shared) & 4.2 & 15.3\% \\
Scorer MLP & 0.004 & 0.013\% \\
LLM prefill ($r{=}0.5$) & 12.8 & 46.7\% \\
LLM decode (10 steps) & 10.4 & 37.9\% \\
\midrule
\textbf{Total} ($r{=}0.5$) & 27.4 & --- \\
\textbf{Total} ($r{=}1.0$) & 41.3 & --- \\
\textbf{Saving} & 13.9 & \textbf{33.6\%} \\
\bottomrule
\end{tabular}
\end{table}

\textbf{Wall-clock}: Variant~B achieves ${\sim}$15\% end-to-end speedup in the 2-pass configuration (pass-1 for scoring still processes full sequence).
In a future 1-pass deployment (scorer integrated into ViT), theoretical speedup reaches 38\%.

% ============================================================================
\section{Discussion}
\label{sec:discussion}

\textbf{Why learned scoring outperforms heuristics}: Training-free methods (FastV, SparseVLM, PruMerge) rely on generic attention patterns or geometric proxies that correlate poorly with driving importance.
Our scorer distills task-relevant signals from the VLA's own internal representations, discovering that driving-critical tokens (lane boundaries, nearby vehicles, traffic signals) receive systematically different importance than generic ``textured'' tokens that attract shallow attention but don't affect trajectory quality.
The +2.84pt gap over random and +0.19pt over the attention teacher confirms that the MLP generalizes beyond its training signal.

\textbf{``Learning what = learning how many''}: Our calibrated scorer + $\tau$-cut shows that a unified scoring mechanism can simultaneously solve token selection (ranking) and budget allocation (counting) without architectural separation.
This contrasts with prior work~\citep{prune2drive2026,mvpruner2026} that treats these as separate problems requiring separate modules.

\textbf{Scaling insight}: Our cross-scale analysis (3B vs.\ 7B) reveals that larger VLAs concentrate vision attention more sharply (95.9\% in top-25\% tokens at 7B vs.\ 91.6\% at 3B), directly explaining why concurrent work reports 50--75\% lossless pruning at 7B scale while we observe 25\% as the lossless threshold at 3B.

\textbf{Limitations}:
(1) The confidence-based fallback (6.6\% of scenes) adds a small overhead; future work will investigate architecturally eliminating catastrophic-failure modes in true token removal.
(2) The 7B cross-model experiment uses nuScenes (different dataset/metric); same-benchmark cross-scale comparison requires a 7B VLA fine-tuned on NAVSIM.
(3) NAVSIM evaluation is open-loop; closed-loop (reactive) evaluation remains future work.

\textbf{From naive RL to Budget RL---reward design matters}:
Our initial RL attempt using raw PDMS (multiplicative composite of 6 sub-scores) as reward failed to surpass SFT (0.889 vs.\ 0.892), because: (1) any sub-score reaching zero collapses the entire reward signal; (2) inter-scene PDMS variance at fixed $r$ is $<$0.01, drowning the advantage in noise.
The breakthrough came from \emph{shaped sub-metric decomposition}: weighting each sub-metric independently (progress$=$0.35, collision$=$0.20, drivable$=$0.15, TTC$=$0.15, direction$=$0.10, comfort$=$0.05) amplifies the gradient signal 10--100$\times$.
Budget RL further adds an efficiency bonus $\beta(1{-}r)$ that incentivizes aggressive pruning when driving quality permits, enabling the scorer to autonomously learn scene-adaptive budget allocation without any manual threshold.

% ============================================================================
\section{Conclusion}
\label{sec:conclusion}

We present \ours{}, a reinforcement-learning-based vision token pruning framework for autonomous driving VLAs.
Our key innovation is a progressive training pipeline---LambdaRank distillation $\to$ shaped RL $\to$ Budget RL---that culminates in a scorer that \emph{autonomously} determines both which tokens to keep and how many per scene, eliminating fixed pruning ratios entirely.
With true token removal and a lightweight confidence-based fallback, \ours{} \textbf{surpasses the unpruned baseline} (PDMS 0.9045 vs.\ 0.8988, Rel.\ 100.6\%) while delivering 15\% wall-clock speedup and 33.6\% FLOPs reduction---demonstrating that removing noisy vision tokens can \emph{improve} driving quality.
The same training method generalizes from 3B to 7B VLAs: a scorer trained from an unfinetuned Qwen-7B backbone works plug-and-play on ImpromptuVLA (nuScenes) without modifying the 7B VLA, confirming that the approach---not a specific set of weights---scales across model sizes.
Our shaped reward design---decomposing driving quality into six independent sub-metric signals---provides a general solution to the reward sparsity challenge in RL for composite-metric optimization, with potential applications beyond token pruning.

% ============================================================================
\bibliography{references}

\end{document}
